Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data

### 2021-04-18 documents/ovando-etal-assessing-global-fisheries.Rmd
cole edits in prep for f and f submission

---
title: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: pnas.csl
output: 
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fish populations play a critical role in tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain. Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added. Estimates of fish populations based primarily on catch histories alone frequently performed roughly as well as a random guess. Obtaining accurate estimates of stock status for the world's exploited fish populations depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

@@

# Significance Statement  {.unnumbered}

<!-- Authors must submit a 120-word-maximum statement about the significance of their research paper written at a level understandable to an undergraduate-educated scientist outside their field of specialty. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. -->

The health of ocean ecosystems and the well-being of millions of people depend on our ability to accurately estimate the state of marine fish populations. However, these assessments typically require extensive amounts of high-quality data that are not available for large numbers of fisheries that are critical sources of economic and nutritional benefits. Many studies have tried to estimate the state of these unassessed fisheries based solely on the history of catches reported for a fishery, but we show here that these catch-based estimates produce unreliable results. We demonstrate that improving our understanding of global fisheries will depend on redoubled efforts at data collection and capacity building.

@@

The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". Marine fisheries are one of the largest anthropogenic impacts in the oceans, that when managed sustainably are a critical source of economic prosperity and food security around the globe. As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries. While our understanding of the state of world fisheries has improved over the last decade, the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their population size relative to reference points [@hilborn2020; @fao2020]. This is a major impediment to ensuring the sustainable development of the world's oceans. In this paper, we assess why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.

What do we currently know about the state of fished marine species around the world? The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries. The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished, 6.2% are underfished, and 34.2% are overfished [@fao2020]. Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012]. The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the state of assessed fish populations around the globe.

That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed". While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020]. While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods. With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals. In addition, the SOFIA process assigns included unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to to the biomass that would on average result in maximized yield over the long-term.

@@

Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide numerical estimates of the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (Free et al. 2020 [@free2020] and references therein). These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being our measure of stock status), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries (e.g. [@cabral2020],[@gaines2018]).

However, Free et al. 2020 [@free2020] demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*. These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model.

Costello et al. 2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in Rosenberg et al. 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions resulting from these catch-only efforts at assessing global fisheries is concerning.

```{r status-tab}

@@

tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
    rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
  select(`FAO Area`,`SOFIA % Overfished`, everything())

 tmp %>% 
  knitr::kable(caption = "Estimates of fish stock status by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating stock status of the individual unassessed fisheries around the world. The types of global assessment efforts detailed in Table.\@ref(tab:status-tab) share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure. Under this assumption, we can obtain better estimates of the state of unassessed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries. However, critical evaluation of the predictions made by these models (Table.\@ref(tab:status-tab)), as well as works such as @pauly2013 and @branch2011 call into question this very idea. In this paper, rather than seeking to improve our knowledge of unassessed fisheries by modifying catch-only models themselves, we ask how much can estimates from catch-only assessments be improved by moving beyond catch-only and augmenting catch histories with additional sources of broadly available fisheries data? We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis, not on incremental improvements to assessment methods based solely on catch histories.

To accomplish this we developed a stock assessment software package, sraplus, whose key attribute is not substantial improvements in model fitting methods per say, but the ability to easily incorporate multiple kinds of fishery data in a statistically rigorous manner. At its most "data-limited", sraplus can work as a catch-only model in the manner of [@froese2017]. These catch histories can then be augmented with prior information on stock status derived from increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases. The FMI database provides self-reported scores reflecting the strength of fisheries management in many fisheries around the world, which sraplus uses to construct informative priors on stock status, e.g. placing a higher probability that fisheries with high FMI scores have better stock status than those with low FMI scores. SAR is a measure of the intensity of trawl fishing within a particular area, which sraplus uses to construct an informative prior on the magnitude of fishing mortality. While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries. As an alternative though, Rousseau et al. 2019 [@rousseau2019b] present a global reconstruction of fishing effort for countries around the world. We pair these effort reconstructions (both nominal effort and effective effort assuming an annual rate of efficiency increase of 2.6%) with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status. In this way then sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).

@@

We then used sraplus to evaluate the performance of estimates of global fisheries derived by using different combinations of these broadly available data, and as a benchmark compared these results with estimates from the increasingly utilized default settings of the catch-only version of the CMSY described in Froese et al. 2017 [@froese2017] (See Table.S1). We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history. *B/B~MSY~* values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data. This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries. We also estimated *B/B~MSY~* values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM itself, as measure of the ability of models like sraplus if given perfect information. We assessed model performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished). This experiment structure allows us to assess the ability of catch-only models to estimate the *B/B~MSY~* values reported in RAM, and evaluate how much these estimates can be improved by augmenting the catch histories with additional sources of information.

We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data at the global scale. We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores. We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al. 2019 [@rousseau2019] to each stock. As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available. We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus. Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values.

While CMSY systemically overestimated fishing mortality rates and underestimated stock status for these case study RAM stocks, use of the fishery-specific SAR, FMI, and regional effort data, produced substantially more accurate results for both *B/B~MSY~* and *F/F~MSY~* (Fig.\@ref(fig:cs-plot)). This is not in and of itself a surprising result, but demonstrates how 1) estimates of stock status based on catch histories alone can be both imprecise and biased and how 2) augmenting catch histories with additional data from the fisheries in question can produce meaningful improvements.

<!-- Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one. CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing. Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~. Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit. This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values (Fig.\@ref(fig:cs-plot)). -->

## Global Value of Information {.unnumbered}

<!-- We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status. We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history. B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RAM stocks with regional-level SAR, FMI, and effort data. This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries. -->

<!-- We also estimated B/B~MSY~ values of our candidate RAM stocks by using sraplus to fit to an abundance index drawn directly from RAM. We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with the CMSY method described in Froese et al. 2017 [@froese2017] (See Table.\@ref(tab:dat-desc)). We assessed performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished). -->

We first consider the performance of sraplus when the model is given a perfect index of abundance from RAM for each stock in RAM it is tasked with assessing. Assessment models in RAM are typically much more structurally complex than the simple biomass dynamics used in sraplus, and so this "perfect information" test tells us how much of a penalty in model performance we are likely to pay due to model misspecification alone. The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RAM were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "RAM Index"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "RAM Index"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "RAM Index"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves. We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status beyond those derived from catch-only methods. Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The status of most stocks in RAM is relatively good, with recent *B/B~MSY~* values generally near one. This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than the RAM abundance indices are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RAM Index","Guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "Guess"])`.

Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process. Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions. For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map). We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)). Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.

knitr::kable(perf, digits = 2,
             caption = "Global performance statistics in the most recent year available of models using different sources of data. mpe = median percent error (bias), mape = median absolute percent error (error), accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished). Performance is judged relative to reported values in RAM Legacy Stock Assessment Database. See Table.5 for details of data types.",
              format = "pipe")

```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources. RAM Index refers to catch and abundance index drawn from RAM. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
ram_mpe_map_plot
```

```{r mape-map, fig.cap = "Median absolute percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources. RAM Index refers to catch and abundance index drawn from RAM. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
ram_mape_map_plot
```

```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources. RAM Index refers to catch and abundance index drawn from RAM. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}

The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RAM, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status. What sources of data might provide the greatest value in improving our estimates of global stock status? We used sraplus together with the RAM database to estimate the value of information, measured as the average reduction in root mean squared error, resulting from use of different kinds of data (Fig.\@ref(fig:voi-plot)). Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available. Interestingly though, having access to only an accurate estimate of F/F~MSY~ in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself. While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average, due to the lack of historical context for recent trends in abundance. We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020].

Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals. Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fish stocks remains uncertain in many parts of the world. The hope of efforts using catch-only models to estimate the status of global unassessed fisheries is fundamentally that we can learn something meaningful about the state of a fished population simply by knowing something about its catch history and life history. Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RAM. While in some cases addition of globally available data such a FMI scores, SAR values, or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)).

>>>>

---
title: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: fish-and-fisheries.csl
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
  bookdown::word_document2:
    reference_docx: template.docx
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fish populations play a critical role in tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the word's 'unassessed' fish stocks remains highly uncertain. Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added. Estimates of fish populations based primarily on catch histories alone on average performed 29% better than a random guess. But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time. Obtaining accurate estimates of stock status for the world's exploited fish populations depends on prioritizing the collection of high-priority data around the globe, rather than the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

@@

<!-- # Significance Statement  {.unnumbered} -->

<!-- <!-- Authors must submit a 120-word-maximum statement about the significance of their research paper written at a level understandable to an undergraduate-educated scientist outside their field of specialty. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. -->

<!-- The health of ocean ecosystems and the well-being of millions of people depend on our ability to accurately estimate the state of marine fish populations. However, these assessments typically require extensive amounts of high-quality data that are not available for large numbers of fisheries that are critical sources of economic and nutritional benefits. Many studies have tried to estimate the state of these unassessed fisheries based solely on the history of catches reported for a fishery, but we show here that these catch-based estimates produce unreliable results. We demonstrate that improving our understanding of global fisheries will depend on redoubled efforts at data collection and capacity building.  -->

@@

The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". Marine fisheries are one of the largest anthropogenic impacts in the oceans aside from the effects of climate change. When managed sustainably fisheries are a critical source of economic prosperity and food security around the globe. As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries. While our understanding of the state of world fisheries has improved over the last decade, the majority of the world's fish populations, making up roughly 50% of marine landings, lack formal statistical assessments (stock assessments) of their population size relative to reference points [@hilborn2020; @fao2020]. This is a major impediment to ensuring the sustainable development of the world's oceans. In this paper, we consider why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.

What do we currently know about the state of fished marine species around the world? The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries. The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished (at or near targets for sustainable food production), 6.2% are underfished, and 34.2% are overfished [@fao2020]. Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments reported in the RAM Legacy Stock Assessment Database (RAM) [@ricard2012]. The "assessed" fisheries in RAM represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the state of assessed fish populations around the globe.

That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed". While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020]. While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods. With the surge in data availability, such as global assessments of management strength, trawl footprints, and fishing effort, the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals. In addition, the SOFIA process assigns unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to to the biomass that would on average result in maximized yield over the long-term.

@@

Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide numerical estimates of the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (@free2020 and references therein). These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being our measure of stock status), from characteristics of a fishery's catch history, and have been the foundation of many subsequent analyses of the future of global fisheries (e.g.  @cabral2020 and @gaines2018).

<!-- These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates of *B/B~MSY~* and other reference points reported in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RAM to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model. -->

<!-- @costello2016a found similar rankings of regions in terms of stock status as RAM, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in @rosenberg2018 demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. Simply comparing catch divided by maximum catch, in the manner of @pauly2007, finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly exploited regions as measured by RAM and the FAO resulting from these catch-only efforts at assessing global fisheries is concerning. -->

```{r status-tab, include=FALSE}

@@

tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RAM`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
    rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
  select(`FAO Area`,`SOFIA % Overfished`,`B/BMSY RAM`, everything())

 tmp %>% 
  knitr::kable(caption = "Estimates of fish stock status by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, Catch / max(Catch)",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

However, @free2020 demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*. While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating stock status of the individual unassessed fisheries around the world. The types of global assessment efforts share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure. Under this assumption, we can obtain better estimates of the state of unassessed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries. However, further evaluation of the predictions made by these models [@free2020;@bouch2021;@pons2020] as well as works such as @pauly2013 and @branch2011 call into question this very idea.

In this paper we ask, how much can estimates from catch-only assessments be improved by moving beyond catch-only and augmenting catch histories with additional sources of broadly available fisheries data? We answer this question by aggregating a set of broadly distributed datasets that might be of use to global fishery assessment. We then used our new assessment method, sraplus, to evaluate the performance status estimates derived from different combinations of these broadly available data. Wecompared these results with estimates from the increasingly utilized (e.g. @palomares2020) default settings of the catch-only version of CMSY described in @froese2017. We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis, not on incremental improvements to assessment methods based solely on catch histories.

# Methods {.unnumbered}

The key attribute of sraplus is not substantial improvements in model fitting methods per say, but the ability to easily incorporate multiple kinds of fishery data potentially used in SRA style analyses in a statistically rigorous manner. We paired these data with fisheries in RAM, and then used sraplus to generate predictions of B/B~MSY~ in the most recent year of each fishery across different combinations of the data types. We then compared the predictions generated by sraplus using different types of data, along with predictions for the same fisheries generated by CMSY, to the reported values in RAM.

All analysis were conducted in the R programming language [@rcoreteam2019]. Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018]. The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries). Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.

## Data Sources {.unnumbered}

At its most "data-limited", sraplus can work as a catch-only model in the manner of @froese2017. These catch histories can then be augmented with prior information on stock status derived from increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases. The FMI database provides self-reported scores reflecting the strength of fisheries management in many fisheries around the world, which sraplus uses to construct informative priors on stock status, e.g. placing a higher probability that fisheries with high FMI scores have better stock status than those with low FMI scores. SAR is a measure of the intensity of trawl fishing within a particular area, which sraplus uses to construct an informative prior on the magnitude of fishing mortality. While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries. As an alternative, @rousseau2019b present a global reconstruction of fishing effort for countries around the world. We pair these effort reconstructions (both nominal effort and effective effort assuming an annual rate of efficiency increase of 2.6%) with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status. In this way sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).

@@

<!-- We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database.  -->
